grammar Lua;

chunk
    : block EOF
    ;

block
    : stat* retstat?
    ;

stat
    : ';'                                                                           # StatEmptySemicolon
    | varlist '=' explist                                                           # StatAssignment
    | functioncall                                                                  # StatFunctionCall
    | 'do' block 'end'                                                              # StatDo
    | 'while' exp 'do' block 'end'                                                  # StatWhile
    | 'repeat' block 'until' exp                                                    # StatRepeat
    | 'if' exp 'then' block ('elseif' exp 'then' block)* ('else' block)? 'end'      # StatIfThenElse
    | 'for' NAME '=' exp ',' exp (',' exp)? 'do' block 'end'                        # StatNumericFor
    | 'for' namelist 'in' explist 'do' block 'end'                                  # StatGenericFor
    | 'function' funcname funcbody                                                  # StatFunction
    ;

attnamelist
    : NAME attrib (',' NAME attrib)*
    ;

attrib
    : ('<' NAME '>')?
    ;

retstat
    : 'return' explist? ';'?                                                        # StatReturn
    ;

funcname
    : NAME ('.' NAME)* (':' NAME)?
    ;

varlist
    : var (',' var)*
    ;

namelist
    : NAME (',' NAME)*
    ;

explist
    : exp (',' exp)*
    ;

exp
    : 'nil'                                                                         # ExpNil
    | 'false'                                                                       # ExpFalse
    | 'true'                                                                        # ExpTrue
    | number                                                                        # ExpNumber
    | string                                                                        # ExpString
    | functiondef                                                                   # ExpFunctionDef
    | prefixexp                                                                     # ExpPrefixExp
    | tableconstructor                                                              # ExpTableConstructor
    | <assoc=right> exp operatorPower exp                                           # ExpOperatorPower
    | operatorUnary exp                                                             # ExpOperatorUnary
    | exp operatorMulDivMod exp                                                     # ExpOperatorMulDivMod
    | exp operatorAddSub exp                                                        # ExpOperatorAddSub
    | <assoc=right> exp operatorStrcat exp                                          # ExpOperatorStrcat
    | exp operatorComparison exp                                                    # ExpOperatorComparison
    | exp operatorAnd exp                                                           # ExpOperatorAnd
    | exp operatorOr exp                                                            # ExpOperatorOr
    ;

prefixexp
    : varOrExp nameAndArgs*
    ;

functioncall
    : varOrExp nameAndArgs+
    ;

varOrExp
    : var | '(' exp ')'
    ;

var
    : (NAME | '(' exp ')' varSuffix) varSuffix*
    ;

varSuffix
    : nameAndArgs* ('[' exp ']' | '.' NAME)
    ;

nameAndArgs
    : (':' NAME)? args
    ;

args
    : '(' explist? ')' | tableconstructor | string
    ;

functiondef
    : 'function' funcbody
    ;

funcbody
    : '(' parlist? ')' block 'end'
    ;

parlist
    : namelist (',' '...')? | '...'
    ;

tableconstructor
    : '{' fieldlist? '}'
    ;

fieldlist
    : field (fieldsep field)* fieldsep?
    ;

field
    : '[' exp ']' '=' exp | NAME '=' exp | exp
    ;

fieldsep
    : ',' | ';'
    ;

operatorOr
	: 'or';

operatorAnd
	: 'and';

operatorComparison
	: '<' | '>' | '<=' | '>=' | '~=' | '==';

operatorStrcat
	: '..';

operatorAddSub
	: '+' | '-';

operatorMulDivMod
	: '*' | '/' | '%' | '//';

operatorUnary
    : 'not' | '#' | '-';

operatorPower
    : '^';

number
    : INT | FLOAT
    ;

string
    : NORMALSTRING
    ;

// LEXER

NAME
    : [a-zA-Z_][a-zA-Z_0-9]*
    ;

NORMALSTRING
    : '"' ( EscapeSequence | ~('\\'|'"') )* '"'
    ;

fragment
NESTED_STR
    : '=' NESTED_STR '='
    | '[' .*? ']'
    ;

INT
    : Digit+
    ;

FLOAT
    : Digit+ '.' Digit* ExponentPart?
    | '.' Digit+ ExponentPart?
    | Digit+ ExponentPart
    ;

fragment
ExponentPart
    : [eE] [+-]? Digit+
    ;

fragment
EscapeSequence
    : '\\' [abfnrtvz"'\\]
    | '\\' '\r'? '\n'
    | DecimalEscape
    | HexEscape
    | UtfEscape
    ;

fragment
DecimalEscape
    : '\\' Digit
    | '\\' Digit Digit
    | '\\' [0-2] Digit Digit
    ;

fragment
HexEscape
    : '\\' 'x' HexDigit HexDigit
    ;

fragment
UtfEscape
    : '\\' 'u{' HexDigit+ '}'
    ;

fragment
Digit
    : [0-9]
    ;

fragment
HexDigit
    : [0-9a-fA-F]
    ;

COMMENT
    : '--[' NESTED_STR ']' -> channel(HIDDEN)
    ;

LINE_COMMENT
    : '--'
    (                                               // --
    | '[' '='*                                      // --[==
    | '[' '='* ~('='|'['|'\r'|'\n') ~('\r'|'\n')*   // --[==AA
    | ~('['|'\r'|'\n') ~('\r'|'\n')*                // --AAA
    ) ('\r\n'|'\r'|'\n'|EOF)
    -> channel(HIDDEN)
    ;

WS
    : [ \t\u000C\r\n]+ -> skip
    ;

SHEBANG
    : '#' '!' ~('\n'|'\r')* -> channel(HIDDEN)
    ;
